---
title: "In the Aeroplane Over the Sea - 20 Years On"
#author: "Mike Tapp <michael.tapp@sharecreative.com>"
date: "2019-10-07"
series:
- Theme
- Hugo
output: html_document
---

```{r setup, include=FALSE}
library(lubridate)
library(jsonlite)
library(dplyr)
library(stringr)
library(tidytext)
library(genius)


id <- "Vt1HzmyEXM3LTYLCVjdRsqfMeY9RQk7_HbV23mA2lO9Ljb8cpztOumE_Re4W6DWU"
secret <- "_0vi6DSlLAhdCHStPP3Omq1_0-_Xihvw4xCMR3R1oSy-_j9GeqKZPOuaZzH6hgql7pLT26I5MNjeNPZDA-qubQ"



```

> "I know that she will live forever. She wont ever die --- *Ghost - Neutral Milk Hotel*"

To mark the twentieth anniversary of one of the most influential albums of all time, I thought I'd explore the sentiment of the crptyic and ethereal lyrics of Jeff Mangum.

```{r data call}
nmh_data <- genius_album(artist = "Neutral Milk Hotel", album = "In the Aeroplane Over the Sea")
nmh_tidy_words <- nmh_data %>% 
  unnest_tokens(word, lyric)
```

Bind the sentiment

```{r bind sentiment}

nmh_sentiment <- nmh_tidy_words %>%
  inner_join(get_sentiments("bing")) %>%
  count(track_title, index = line, sentiment) %>%
  tidyr::spread(sentiment, n, fill = 0) %>%
  mutate(sentiment = positive - negative)
```

Plot over the course of the album

```{r plot}
library(ggthemes)
library(ggplot2)
ggplot(nmh_sentiment, aes(index, sentiment, fill = track_title)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~track_title, ncol = 2, scales = "free_x") +
  theme_minimal() +
  labs(title = "Sentiment by Line of each track from Sun Kil Moon's album 'Benji'",
                caption = "Bing sentiment binded to lyrics scraped from Genius",
                x = "line number",
                y = "sentiment") +
  theme(text = element_text(family="Arial", size=12),
        strip.background = element_rect(colour = "#F4F4F4", fill = "#F4F4F4"),
        strip.text.x = element_text(colour = "grey50", face = "bold"))+
  scale_fill_ptol()
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
